### 2025-11-13 draft_group-6.Rmd
draft outline

---
title: "draft_group-6"
output: pdf_document
---

*Paper (40%)* A paper detailing the motivation, methodology, and results of your project, complete with references. (https://dsci445-csu.github.io/project/)

# Outline

## motivation
(10% of the content)\
The Bank Marketing dataset contains information collected from a Portuguese banking institution during multiple direct telemarketing campaigns (source1). *Why??* (We want to justify the efforts of modeling this data)

**Description of dataset** \
Each row represents an individual client that was contacted in an attempt to sell a long-term deposit product. The dataset includes 16 predictor variables describing client demographics, financial attributes, and details about the marketing campaign itself (such as: contact method, call duration, number of contacts, outcome of previous campaigns).
These variables are a mix of categorical and quantitative data, which makes this dataset well-suited for comparing multiple classification models. The target variable indicates whether the client subscribed to a term deposit (yes/no). As usual, the majority of clients do not subscribe (source 2). This causes an imbalance in our data by increasing modeling bias.

**Problem statement/Assumptions** \
We can go over what we are focusing on and what factors we may ignore.

We would like to maintain interpretability in our model. Emphasis is predicated for the model to provide actionable insght to the manager. After all humans love stories.

## methodology 
(60%)

- Data wrangling/cleaning
- For each model (Logistic, LDA, QDA, KNN, Boosting )
  - Model Formulation (training data, with k-fold CV)
  - Feature Selection and model tuning (subset, regularization, dimension reduction)

## results
(25%)

- Model comparison and assessment (Test error, confusion matrix, AUC, BIC, etc.)
- Discussion (make inferrences, determine the favored model)

## references
(5%)