### 2025-11-18 model_fit.Rmd

---
title: "model_fit"
output: html_document
---

load clean data

```{r}

```

## We plan to use the k-fold cross validation approach. 
For consideration: Are we able to make our training sets a sort of proper stratification of our population dataset (Imbalance in amounts of customers subscribed).

```{r}
bank_train <- 
bank_test <- 
```

## Lgoistic Regression

```{r}
library(glmnet)
library(ggplot2)

bank_glm <- glm(y ~., data = bank_train, family = binomial)
summary(bank_glm)
```

estimate test error

## LDA and QDA

```{r}
library(MASS)

bank_lda <- lda(y ~., data = bank_full, subset = train)
bank_lda

bank_qda <- qda(y ~., data = bank_full, subset = train)
```

estimate test error

## KNN
Interpretable in a different way... Can be used to identify customers similar to those that are subscirbed.

```{r}
library(class)

bank_knn <- knn()
```

estimate test error

Optimize K value (w.r.t cross validation test error)

## GAM

```{r}
library(gam)
bank_gam <- gam(y ~ , data = bank_train) # basis functions

```

## Boosting Decision Trees
require post-hoc explanation

```{r}

```

Potential model Mixed effects model (attribute set variables to have a random effect)