### 2025-12-03 draft_group-6.Rmd

---
title: "draft_group-6"
output:
  html_document:
    df_print: paged
---

*Paper (40%)* A paper detailing the motivation, methodology, and results of your project, complete with references. (https://dsci445-csu.github.io/project/)

# Outline

## motivation
(13% of the content)\
The Bank Marketing dataset contains information collected from a Portuguese banking institution during multiple direct telemarketing campaigns (source1). *Why??* (We want to justify the efforts of modeling this data)

**Description of dataset** \
Each row represents an individual client that was contacted in an attempt to sell a long-term deposit product. The dataset includes 16 predictor variables describing client demographics, financial attributes, and details about the marketing campaign itself (such as: contact method, call duration, number of contacts, outcome of previous campaigns).
These variables are a mix of categorical and quantitative data, which makes this dataset well-suited for comparing multiple classification models. The target variable indicates whether the client subscribed to a term deposit (yes/no). As usual, the majority of clients do not subscribe (source 2). This causes an imbalance in our data by increasing modeling bias.

**Problem statement/Assumptions** \
We can go over what we are focusing on and what factors we may ignore.

We would like to maintain interpretability in our model. Emphasis is predicated for the model to provide actionable insght to the manager. After all humans love stories.

## methodology 
(60%)

- Data wrangling/cleaning
- For each model (Logistic, LDA, QDA, KNN, GAM, Boosting)
  - Model Formulation (training data, with k-fold CV)
  - Feature Selection and model tuning (subset, regularization, dimension reduction)

## 3. Data and Preprocessing
3.1 Dataset Description

Source: UCI Machine Learning Repository.

Number of observations and predictors.

Summary of variable types (categorical, numeric, campaign history).

3.3 Exploratory Data Analysis

Distribution of the response variable (class imbalance).

Univariate summaries: histograms, bar plots.

Key relationships between predictors and response.

## 4. Methodology
4.1 Overview of Statistical Learning Framework

Supervised binary classification.

Tradeoff between interpretability and predictive power.

4.2 Resampling Methods (ISLR Ch. 5)
4.2.1 Train/Test Split

Rationale for honest evaluation.

4.2.2 k-Fold Cross-Validation

Applied to Logistic Regression, KNN, and GAM (if included).

Used for model tuning and error estimation.

4.2.3 Bootstrap (Optional)

Estimating coefficient variability in Logistic Regression.

4.3 Candidate Models (ISLR Ch. 4 & 6)

Logistic Regression

Linear Discriminant Analysis (LDA)

Quadratic Discriminant Analysis (QDA)

K-Nearest Neighbors (KNN) with CV-selected k

Generalized Additive Models (GAM) (optional but interpretable)

4.4 Assessment Metrics (ISLR Ch. 6)

Test misclassification rate.

Sensitivity & specificity.

Confusion matrix.

ROC curve and AUC.

BIC (when applicable).

## 5. Results
5.1 Cross-Validation Findings

Optimal tuning parameters (e.g., best k for KNN).

CV-estimated test errors for each model.

5.2 Test Set Performance

Summary table: accuracy, sensitivity, specificity, AUC.

ROC curves for model comparison.

5.3 Interpretation of Key Predictors

Coefficient signs and magnitudes (Logistic).

Group separation interpretation (LDA/QDA).

Partial dependence/smooth terms (GAM).

## 6. Discussion
- Which model offers the best predictive accuracy?
- Which model is most interpretable for a marketing manager?
- How did resampling methods influence the final choice?
- Effects of class imbalance.
- Limitations and suggested future modeling directions.

Future directions: oversampling methods, modern ML models.