### 2025-12-08 plots.Rmd
eda slides

---
title: "plots"
author: "Group 6"
date: "2025-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

# EDA plots as images

# 1. Class imbalance plot (from page 3 of your PDF)
png("imbalance_plot.png", width = 800, height = 500)
ggplot(bank_full, aes(x = y, fill = y)) +
  geom_bar() +
  labs(title = "Distribution of Term Deposit Subscription",
       x = "Subscribed to Term Deposit?",
       y = "Count") +
  theme_minimal(base_size = 14)
dev.off()

# 2. Contact method plot (from your contact type analysis)
png("contact_plot.png", width = 800, height = 500)
ggplot(bank_full, aes(x = contact, fill = y)) +
  geom_bar(position = "fill") +
  labs(title = "Subscription Rate by Contact Type",
       x = "Contact Method",
       y = "Proportion") +
  theme_minimal(base_size = 14)
dev.off()

# 3. Job type plot
png("job_plot.png", width = 800, height = 500)
ggplot(bank_full, aes(x = job, fill = y)) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(title = "Subscription Rate by Job",
       y = "Proportion") +
  theme_minimal(base_size = 12)
dev.off()

# 4. Education plot
png("education_plot.png", width = 800, height = 500)
ggplot(bank_full, aes(x = education, fill = y)) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(title = "Subscription Rate by Education Level",
       y = "Proportion") +
  theme_minimal(base_size = 14)
dev.off()

# 5. Previous outcome plot
png("poutcome_plot.png", width = 800, height = 500)
ggplot(bank_full, aes(x = poutcome, fill = y)) +
  geom_bar(position = "fill") +
  labs(title = "Subscription Rate by Previous Campaign Outcome",
       y = "Proportion") +
  theme_minimal(base_size = 14)
dev.off()

# 6. Campaign contacts plot
png("campaign_plot.png", width = 800, height = 500)
bank_full %>%
  group_by(campaign) %>%
  summarise(success_rate = mean(y == "yes")) %>%
  ggplot(aes(x = campaign, y = success_rate)) +
  geom_line() +
  geom_point() +
  labs(title = "Subscription Success Rate vs. Number of Campaign Contacts",
       x = "Number of Contacts During This Campaign",
       y = "Success Rate") +
  theme_minimal(base_size = 14)
dev.off()

cat("EDA plots saved as PNG files in your current directory!\n")
cat("Files created:\n")
cat("- imbalance_plot.png\n")
cat("- contact_plot.png\n")
cat("- job_plot.png\n")
cat("- education_plot.png\n")
cat("- poutcome_plot.png\n")
cat("- campaign_plot.png\n")

```